\documentclass[../textbook.tex]{subfiles}

\begin{document}

\section{Infinite Groups}

All of the groups we've seen so far are finite in size. We can also construct groups of infinite size; these are known simply as \textbf{infinite groups}.

A quick review: \textit{iso}- means the same and -\textit{morphic} means form. Two groups are considered isomorphic if there exists a one-to-one correspondence taking each element of the first group to an element of the second group---and vice versa---so that the products of the elements map in the same way. Essentially, by renaming the elements of one group, the two groups appear identical. Isomorphic groups have the same structure and size, and thus the group table's structure is also preserved.

\begin{enumerate}
\item Where have you come across the roots \textit{iso}- and -\textit{morphic} before?
\item Can two groups be isomorphic if they have different orders?
\item The rotation group of the regular hexagon, also known as the cyclic group of order $6$, $C_6$, has six elements: the identity, and rotations of $\frac{\pi}{3}$, $\frac{2\pi}{3}$, $\pi$, $\frac{4\pi}{3}$, $\frac{5\pi}{3}$ radians. A rotation of $\frac{\pi}{3}$ generates the group.
\begin{enumerate}
\item Which other rotation can generate the group on its own?
\item What is the period of each element?
\end{enumerate}
\item $C_6$ has the same number of elements as the dihedral group $D_3$.
\begin{enumerate}
\item Are the two groups isomorphic? How do you know?
\item What is the period of each element of $D_3$?
\item What can you say if the set of the periods of each group's elements are not the same?
\item Which subgroups of the cyclic group $C_6$ and $D_3$ are isomorphic?
\end{enumerate}
\item Can an infinite group and a finite group be isomorphic?
\item Do you think all infinite groups are isomorphic to each other? Find a counterexample if you can.
\setcounter{problem_i}{\value{enumi}}
\end{enumerate}

\noindent If an infinite group was somehow ``bigger'' than another, they wouldn't be isomorphic.\footnote{After reading this section, can you think of an example of two infinite groups that aren't isomorphic?} This raises the question: are all infinities equally big?

We can formalize the notion of sizes of infinity. Let's say that two infinite sets are of the same size if their elements can be put into a one-to-one correspondence with each other. For example, the natural numbers $\{1,2,...\}=\mathbb{N}$ and negative integers $\{-1,-2,...\}=-\mathbb{N}$ are of the same size because we have the one-to-one correspondence $\mathbb{N} \ni n\leftrightarrow -n \in -\mathbb{N}$. $1$ is paired with $-1$, $2$ is paired with $-2$, $3$ is paired with $-3$, and so on. Every element of the positive integers has exactly one ``partner'' in the negative integers, and vice versa.

\begin{enumerate}
\setcounter{enumi}{\value{problem_i}}
\item Make guesses to the relative sizes of the following pairs of sets. You may use shorthand like $|a| < |b|$, $|a| > |b|$, $|a| = |b|$, where $|S|$ denotes the order of $S$. After you have made your guesses, we will analyze some of the cases and you can find out how good your intuition was.~\label{prob:infinity_guesses}

\begin{tabular}{llll} % used instead of enumerate to facilitate alignment of the "versus" symbols. There is probably other ways to do this
(a) & natural numbers, $\mathbb{N}$ & vs. & positive even numbers, $2\mathbb{N}$ \\
(b) & natural numbers, $\mathbb{N}$ & vs. & positive rational numbers, $\mathbb{Q}^+$ \\
(c) & natural numbers, $\mathbb{N}$ & vs. & real numbers between zero and one, $[0,1)$ \\
(d) & real numbers, $\mathbb{R}$ & vs. & complex numbers, $\mathbb{C}$ \\
(e) & real numbers, $\mathbb{R}$ & vs. & points on a line \\
(f) & points on a line & vs. & points on a line segment \\
(g) & points on a line & vs. & points on a plane \\
(h) & rational numbers, $\mathbb{Q}$ & vs. & Cantor set\footnotemark
\end{tabular}
\footnotetext{Consult an online source for the definition of the Cantor set.}

\setcounter{problem_i}{\value{enumi}}
\end{enumerate}

\noindent It turns out that studying infinity involves some strange mathematics. For instance, even though it seems that there should be half as many positive even numbers as natural numbers (see 7a), we can construct a one-to-one correspondence between the two sets, such that every natural number $n$ is paired with a positive even integer $2n$ and vice versa. The existence of this correspondence means that the two sets are equal in size. In symbols, $$(2\mathbb{N} \ni 2n \leftrightarrow n \in \mathbb{N}) \leftrightarrow |2\mathbb{N}|=|\mathbb{N}|.$$ % "bijection" removed because it was only mentioned once. \ni is the set \in symbol but backwards lol

\noindent More surprisingly, we can establish a one-to-one correspondence between the non-negative rational numbers $\mathbb{Q}_{\geq 0}$ and the natural numbers. Draw the numbers $\mathbb{Q}_{\geq 0}$ in a grid as shown in Figure~\ref{fig:rat_ordering_1}, and pair these numbers up with the numbers $1,2,3,...$ in the pattern shown with the arrows. You will eventually list all of the non-negative rational numbers, multiple times, into a correspondence with the natural numbers. To make it
one-to-one, only pair the rational numbers that are in simplest form. For example, we pair $2$ with $\frac{1}{1}$ instead of $\frac{0}{2}$, since $\frac{0}{1}$ is the same number and is already paired with $1$. This method prevents multiple natural numbers from being paired up with the same rational number: the correspondence is now one-to-one, and is depicted in Figure~\ref{fig:rat_ordering_2}.

The real numbers between $0$ and $1$, however, cannot be put into a one-to-one correspondence with $\mathbb{N}$ (see 7c). We will prove this by contradiction. Suppose I told you that I have paired each real number $0 \leq r_k < 1$ with a unique natural number $k$, and vice versa. Then, you can construct a real number $r_\omega$ whose $1$\textsuperscript{st} digit (after the decimal point) differs from $r_1$’s, whose $2$\textsuperscript{nd} digit differs from $r_2$’s, and so on. In other words, it differs from $r_n$ in the $n$\textsuperscript{th} digit. We can make better sense of this construction by writing the numbers out in a table, as shown in Figure~\ref{fig:cantor_diag}. Your new number $r_\omega$ is at the bottom; it differs from all the previous numbers in at least one place, so it is a new real number. Therefore, my original list is incomplete, and such a correspondence doesn’t exist. This is called a \textbf{diagonalization argument} because the differing digits make a diagonal. % mfw CHAMPERNOWNE CONSTANT and R OMEGA

% I did not do these by hand; they were generated by accessories/scripts/create_rational_correspondence_table.py
% This command makes the cancelling symbol for the rational <-> natural number correspondence
\newcommand{\corr}[3] {$\cancelto{#3}{\frac{#1}{#2}}$}
% This command colors stuff gray
\newcommand{\tcg}[1] {\textcolor{gray}{$#1$}}

% Figure: Correspondence between rationals and naturals
\begin{figure}[h]
	\begin{center}
		\begin{minipage}[b]{0.45\textwidth}
			\centering
			\renewcommand*{\arraystretch}{2} % stops fractions from colliding by putting more vertical stretch
			\begin{tabular}{lllllllll} % 9 columns
				 & & \multicolumn{5}{c}{Num. $\rightarrow$} & & \\ % "Num. ->" on the top
				 & & $0$ & $1$ & $2$ & $3$ & $4$ & $5$ & $\cdots{}$ \\
				\multirow{5}{*}{\rotatebox[origin=c]{-90}{Den. $\rightarrow$}}  % "Den. ->" on the left, rotated sideways

				 & $1$ & \corr{0}{1}{1} & \corr{1}{1}{3} & \corr{2}{1}{6} & \corr{3}{1}{10} & \corr{4}{1}{15} & \corr{5}{1}{21} & $\cdots{}$ \\
				 & $2$ & \corr{0}{2}{2} & \corr{1}{2}{5} & \corr{2}{2}{9} & \corr{3}{2}{14} & \corr{4}{2}{20} & \corr{5}{2}{27} & $\cdots{}$ \\
				 & $3$ & \corr{0}{3}{4} & \corr{1}{3}{8} & \corr{2}{3}{13} & \corr{3}{3}{19} & \corr{4}{3}{26} & \corr{5}{3}{34} & $\cdots{}$ \\
				 & $4$ & \corr{0}{4}{7} & \corr{1}{4}{12} & \corr{2}{4}{18} & \corr{3}{4}{25} & \corr{4}{4}{33} & \corr{5}{4}{42} & $\cdots{}$ \\
				 & $5$ & \corr{0}{5}{11} & \corr{1}{5}{17} & \corr{2}{5}{24} & \corr{3}{5}{32} & \corr{4}{5}{41} & \corr{5}{5}{51} & $\cdots{}$ \\
				& $\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ & $\ddots{}$ \\

				% cdots = centered 3 dots, vdots = vertical 3 dots, ddots = 3 dots going down-right
			\end{tabular}
			\vspace*{0.5\baselineskip}
		\end{minipage}
		\hfill % put a little padding between the minipages
		\begin{minipage}[b]{0.45\textwidth}
			\centering
			\renewcommand*{\arraystretch}{2}
			\begin{tabular}{lllllllll} % 9 columns
				 & & \multicolumn{5}{c}{Num. $\rightarrow$} & & \\
				 & & $0$ & $1$ & $2$ & $3$ & $4$ & $5$ & $\cdots{}$ \\
				\multirow{5}{*}{\rotatebox[origin=c]{-90}{Den. $\rightarrow$}}

				 & $1$ & \corr{0}{1}{1} & \corr{1}{1}{2} & \corr{2}{1}{4} & \corr{3}{1}{6} & \corr{4}{1}{10} & \corr{5}{1}{12} & $\cdots{}$ \\
				 & $2$ & \textcolor{gray}{\corr{0}{2}{\textcolor{black}{(1)}}} & \corr{1}{2}{3} & \textcolor{gray}{\corr{2}{2}{\textcolor{black}{(2)}}} & \corr{3}{2}{9} & \textcolor{gray}{\corr{4}{2}{\textcolor{black}{(4)}}} & \corr{5}{2}{17} & $\cdots{}$ \\
				 & $3$ & \textcolor{gray}{\corr{0}{3}{\textcolor{black}{(1)}}} & \corr{1}{3}{5} & \corr{2}{3}{8} & \textcolor{gray}{\corr{3}{3}{\textcolor{black}{(2)}}} & \corr{4}{3}{16} & \corr{5}{3}{21} & $\cdots{}$ \\
				 & $4$ & \textcolor{gray}{\corr{0}{4}{\textcolor{black}{(1)}}} & \corr{1}{4}{7} & \textcolor{gray}{\corr{2}{4}{\textcolor{black}{(3)}}} & \corr{3}{4}{15} & \textcolor{gray}{\corr{4}{4}{\textcolor{black}{(2)}}} & \corr{5}{4}{26} & $\cdots{}$ \\
				 & $5$ & \textcolor{gray}{\corr{0}{5}{\textcolor{black}{(1)}}} & \corr{1}{5}{11} & \corr{2}{5}{14} & \corr{3}{5}{20} & \corr{4}{5}{25} & \textcolor{gray}{\corr{5}{5}{\textcolor{black}{(2)}}} & $\cdots{}$ \\
				& $\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ &$\vdots{}$ & $\ddots{}$ \\
			\end{tabular}
			\vspace*{0.5\baselineskip}
		\end{minipage}
	\end{center}
	\vspace*{-2\baselineskip}
	\begin{center}
		\begin{minipage}[t]{0.45\textwidth}
			\caption{A correspondence between $\mathbb{Q}^+$ and $\mathbb{N}$, but not one-to-one.}
			\label{fig:rat_ordering_1}
		\end{minipage}
		\hfill % again, put a little padding
		\begin{minipage}[t]{0.45\textwidth}
			\caption{The previous correspondence, with duplicates parenthesized but not counted; it is now one-to-one.}
			\label{fig:rat_ordering_2}
		\end{minipage}
	\end{center}

	\begin{center}
		\begin{minipage}[b]{\textwidth}
			\setlength{\tabcolsep}{3.5pt} % Decreases the horizontal padding, which is useful here because we only have single digits
			\centering
			\begin{tabular}{rlllllllllllllllllll} % 19 columns
				& & & \multicolumn{16}{c}{Decimal Expansion $\rightarrow$} \\
				% A note: the \Aboxed command from math tools allows boxed things to be aligned from inside the box. Without this, the entry would be aligned from the left side of the BOX, not the DIGIT, which is a bit triggering
				$r_{1}$ & $0$ & \tcg{0.}\Aboxed{ & 0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & $\cdots{}$ \\
				$r_{2}$ & $\frac{1}{4}$ & \tcg{0.} & \tcg{2} \Aboxed{ & 5} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & \tcg{0} & $\cdots{}$ \\
				$r_{3}$ & $\frac{\pi}{4}$ & \tcg{0.} & \tcg{7} & \tcg{8} \Aboxed{ & 5} & \tcg{3} & \tcg{9} & \tcg{8} & \tcg{1} & \tcg{6} & \tcg{3} & \tcg{3} & \tcg{9} & \tcg{7} & \tcg{4} & \tcg{4} & \tcg{8} & \tcg{3} & $\cdots{}$ \\
				$r_{4}$ & $C_{10}$ & \tcg{0.} & \tcg{1} & \tcg{2} & \tcg{3} \Aboxed{ & 4} & \tcg{5} & \tcg{6} & \tcg{7} & \tcg{8} & \tcg{9} & \tcg{1} & \tcg{0} & \tcg{1} & \tcg{1} & \tcg{1} & \tcg{2} & \tcg{1} & $\cdots{}$ \\
				$r_{5}$ & $e-2$ & \tcg{0.} & \tcg{7} & \tcg{1} & \tcg{8} & \tcg{2} \Aboxed{ & 8} & \tcg{1} & \tcg{8} & \tcg{2} & \tcg{8} & \tcg{4} & \tcg{5} & \tcg{9} & \tcg{0} & \tcg{4} & \tcg{5} & \tcg{2} & $\cdots{}$ \\
				$r_{6}$ & $\frac{\sqrt{2}}{10}$ & \tcg{0.} & \tcg{1} & \tcg{4} & \tcg{1} & \tcg{4} & \tcg{2} \Aboxed{ & 1} & \tcg{3} & \tcg{5} & \tcg{6} & \tcg{2} & \tcg{3} & \tcg{7} & \tcg{3} & \tcg{0} & \tcg{9} & \tcg{5} & $\cdots{}$ \\
				$\vdots{}$ & $\vdots{}$ ~~ & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ &  & $\vdots{}$ & $\ddots{}$ \\
				$r_\omega$ &  & $0.$ & $1$ & $6$ & $6$ & $5$ & $9$ & $2$ & \multicolumn{5}{l}{$\cdots{}$} \\
			\end{tabular}
			\vspace*{0.5\baselineskip}
		\end{minipage}
	\end{center}
	\vspace*{-2\baselineskip}
	\begin{center}
		\begin{minipage}[t]{\textwidth}
			\caption{Cantor's diagonal argument. Notice how, by construction, $r_\omega$ differs from $r_i$ in the circled digits.}
			\label{fig:cantor_diag}
		\end{minipage}
	\end{center}
	\vspace*{-2\baselineskip}
\end{figure}

\begin{enumerate}
\setcounter{enumi}{\value{problem_i}}
\item Now, please return to Problem~\ref{prob:infinity_guesses} and revise your answers. Justify each answer by producing a one-to-one correspondence, or showing the impossibility of doing so. Part (h) is an optional challenge.
\setcounter{problem_i}{\value{enumi}}
\end{enumerate}

\noindent The infinities in Problem~\ref{prob:infinity_guesses} come in only two sizes: \textbf{countable} infinity, like the number of natural numbers; and \textbf{uncountable} infinity, like the number of real numbers. There are, in fact, an infinite number of sizes of infinity, but these two main types are the only ones we'll deal with in this class.\footnote{Are there any infinities \textit{between} the two we’ve discussed? This is a very deep mathematical question, known as the continuum hypothesis. It turns out that both the answer ``yes'' and ``no'' are consistent with the rest of our mathematics, so either can be taken as an axiom.}

Two infinite groups can be the same infinite size and still not be isomorphic, in the same way that two finite groups of the same size are sometimes not isomorphic (like $D_3\neq S_6$). For example, the group of all rotations of a rational number of degrees between $0^\circ$ and $360^\circ$ about the origin is countably infinite. So is the group of integers---positive and negative---under addition. But these two groups have completely different structures. For example, the former has two
elements which are their own inverse: $0^\circ$ and $180^\circ$. The latter has only one such element: $0^\circ$.

\begin{enumerate}
\setcounter{enumi}{\value{problem_i}}
\item Here’s a list of infinite sets, each with an operation. For each pair, answer: (i)~Does it form a group? (ii)~Which previous group(s) is it isomorphic to, if any?
\begin{enumerate}
\begin{multicols}{2}\raggedcolumns
\item natural numbers, addition
\item integers, addition
\item even integers, addition
\item odd integers, addition
\item rational numbers, addition
\item real numbers, addition
\item complex numbers, addition
\item integers, multiplication
\item integer powers of $2$, multiplication
\item rational numbers, multiplication
\item rational numbers excluding $0$, multiplication
\item real numbers excluding $0$, multiplication
\item complex numbers, multiplication
\item rotation by a rational number of degrees
\item rotation by a rational number of radians
\item rotation by an integer number of radians
\end{multicols}
\end{enumerate}
\item Can an irrational number taken to an irrational power ever be rational? Consider the potential example $a = \sqrt{2}^{\sqrt{2}}$. To help you answer this question, let $b = a^{\sqrt{2}}$. Simplify $b$, and explain why we don’t need to know whether $a$ is rational or irrational.
\end{enumerate}


\end{document}
